---
title: "Penguins size compnarisons"
author: "Satie Airame"
date: "11/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize) # contains Cohen's D effect size function
library(broom)
library(ggbeeswarm)
library(car)
```
# Part 1. T-tests

Mass comparison: male and female chinstrap
```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap")

# Histograms

ggplot(data = chinstraps, aes(x = body_mass_g)) +
  geom_histogram(bins = 10) +
  facet_wrap(~sex)

## QQ Plot 

ggplot(data = chinstraps, aes(sample = body_mass_g))+
  geom_qq()+
  facet_wrap(~sex)

## Get descriptive statistics

chinstraps %>% 
  group_by(sex) %>% 
  summarize(
    mean_mass = mean(body_mass_g),
    median_mass = median(body_mass_g),
    sd_mass = sd(body_mass_g),
    sample_size = n()
  )

```

Create vectors of body_mass_g for female and male chinstrap penguins separately (in preparation for running two-sample t-test)

```{r}
chinstraps_f <- chinstraps %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g) # pull the values from a function that we specify

chinstraps_m <- chinstraps %>% 
  filter(sex == "chinsmale") %>% 
  pull(body_mass_g)
```

## T-tests, a couple different ways

1. Is there a significant difference in body mass between male and female penguins?

```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest

cohen.d(chinstraps_f, chinstraps_m)
```

## Lowest line is the mean of the two groups in the order that we input them.
## t-value is how different from 0means are on the t-distribution, df = measure of sample size, p-value is probability that we got sample means at least as different as this by random chance if they were drawn from populations of the same mean.
## conclusion is that there is a significant different in mean body mass between male and female chinstrap penguins.
## Cohen's D = -1.26 is large so there is a big difference between male and female chinstrap penguins

Use `broom::tidy()` to get messy model outputs into tidy data frames that are easy to reference! 

```{r}
chinstrap_out <- broom::tidy(chinstrap_ttest)
chinstrap_out
```

You can **reference values from model outputs in your R markdown text**:

t(`r round(chinstrap_out$parameter,2)`) = `r round(chinstrap_out$statistic,2)`, *p* = `r chinstrap_out$p.value`.

Are male chinstrap penguins significantly larger (body mass) than female chinstrap penguins? (One-sided test, only interested in one tail of t-distribution), order matters!

```{r}
t.test(chinstraps_m, chinstraps_f, alternative = "greater") # or "less" depending on argument and order. Is male mass greater than female mass?
```

## Male chinstrap penguins body mass is sigificantly greater than female chinstrap penguin body mass.

## One-way (single factor with discrete groups) ANOVA

## 1. Table of counts by species and sex
```{r}
penguin_counts <- penguins %>% 
  count(species, sex)
penguin_counts
```
## Over 30 samples so likely this will be normally distributed but do exploratory analysis and distribution.

## Create a subset of only female penguins, explore normality

```{r}
penguin_f <- penguins %>% 
  filter(sex == "female")

ggplot(data = penguin_f, aes(x = flipper_length_mm))+
  geom_histogram(aes(fill = species))+
  facet_wrap(~species, scales  = "free") # without scales = "free" then the x and y axes are exactly the same to more easily compare
ggplot(data = penguin_f, aes(sample = flipper_length_mm))+
  geom_qq(aes(color = species))+
  facet_wrap(~species, scales  = "free")
```

## Summary statistics

Starting from penuins_f, create a summary table that groups by species and returns the following: mean of flipper length, sd of flipper length, sample size of flipper length, variance of flipper length "var", standard error "sd/sqrt(n)"

```{r}
penguin_f_summary <- penguin_f %>% 
  group_by(species) %>% 
  summarize(
    mean_flipper_length = mean(flipper_length_mm),
    sd_flipper_length = sd(flipper_length_mm),
    sample_size_flipper_length = n(),
    se_flipper_length = sd(flipper_length_mm)/sqrt(n()),
    variance_flipper_length =var(flipper_length_mm)
  )

leveneTest(flipper_length_mm ~ species, data = penguin_f) # leveneTest in car package
```

### Beeswarm plot with mean +- 1sd

```{r}
ggplot() + 
  geom_beeswarm(data = penguin_f, 
                aes(x = species, y = flipper_length_mm),
                color = "gray70",
                size = 3,
                alpha = 0.5) +
  geom_point(data = penguin_f_summary, aes(x = species, y = mean_flipper_length),
             size = 4) +
  geom_errorbar(data = penguin_f_summary, 
                aes(x = species,
                    ymax = mean_flipper_length + sd_flipper_length,
                    ymin = mean_flipper_length - sd_flipper_length),
                width = 0.1) +
  theme_light()
```

### One-way ANOVA
aov function runs one-way anova,
unequal sample sizes = do type three anova

```{r}
penguin_aov <- aov(flipper_length_mm ~ species, data = penguin_f)
summary(penguin_aov)

TukeyHSD(penguin_aov)

```

F statistic is ratio between group variance / within group variance
F = 411 is large so there is a small probability that we could have gotten sample means that were at least this different if we had drawn sample means from populations with the same mean flipper length. Reject null hypothesis that means of all groups are the same.

Tukey's HSD shows there is significant difference between all the pairwise combinations.

If unbalanced design (sample sizes, groups, distributions), then probably want to do Type III Anova, orthogonal 

Type III Anova with car package:

```{r}
penguin_lm <- lm(flipper_length_mm ~ species, data = penguin_f)
penguin_aov_t3 <- Anova(penguin_lm, type = "III")
penguin_aov_t3
```







